---
title: "Percent change"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r load}
library(tidyverse)
library(iml)
library(ranger)
load("data/ES-RF-all.RDA")
load("data/ES-RF-delta.RDA")
```

```{r metric_cor}

cor.met <- corrplot::corrplot(cor(dat[,17:42]),
                   order = "hclust", hclust.method = "complete",addrect = 2,
                   # type = "upper", 
                   tl.col = 'black', tl.srt = 55, tl.cex = 0.8, tl.pos = "lt",
                   number.cex = 0.5, number.digits=1)

# cor.met$arg$type <- "full"

# corrplot::corrRect.hclust(cor.met, k = 2, method = "complete")

cor.met <- corrplot::corrplot.mixed(cor(dat[,17:42]),
                   order = "hclust", hclust.method = "complete",addrect = 2,
                   # type = "upper", 
                   tl.col = 'black', tl.srt = 55, tl.cex = 0.8, tl.pos = "lt",
                   number.cex = 0.6, number.digits=2, number.col="black")

```

## Ecosystem service improvements depend on the baseline condition

```{r data_setup}


dat.long <- dat %>% 
  pivot_longer(cols = c(1:16), names_to = "ES_response", values_to = "ES_value") %>%
  pivot_longer(cols = c("ntr_pland", "ann_agr_pland"), names_to = "Landuse",
               values_to = "Landuse_value") %>%
  select(ES_response, ES_value, Landuse, Landuse_value) %>%
  mutate(ES = gsub("delta_|_base|_(incr|red).margin|pc_|_max.margin","", ES_response)) %>%
  mutate(ES = case_match(ES,
                         "sed" ~ "sed_export",
                         "pol" ~ "pollinators",
                         "nexp" ~ "n_export",
                         .default = ES)) %>%
  mutate(response_type =
           gsub("sed|export|n_|pollinators|qb|_|pol|nexp|red.|incr.|max","",ES_response)) %>%
  # mutate(response_type = case_match(
  #   response_type,
  #   "base" ~ "Baseline ES [/ha]",
  #   "delta" ~ "\u0394 ES [/ha]",
  #   "pcmargin" ~ "% \u0394 ES",
  #   "delta.margin" ~ "\u0394 ES per ha margin"
  # )) %>%
  mutate(response_type = factor(response_type, 
                                levels = c("base",
                                           "delta",
                                           "pcmargin",
                                           "delta.margin")),
         ES = factor(ES, levels = c("pollinators","qb", "sed_export", "n_export")))

```

Ecosystem service improvements depend on the baseline condition, except for pollination.

This pattern holds even if we factor out margin area. 

The trend is different if we look at % change in ecosystem services from baseline, per margin area

```{r percentchangedemo, message=FALSE, fig.width = 8, fig.height=9}

# dat_wider <- dat.long %>% select(-ES_response) %>% 
#   pivot_wider(names_from = response_type, values_from = ES_value)

# lu.v.base <- ggplot(dat_wider %>% filter(Landuse=="ntr_pland"), 
#                        aes(Landuse_value,`Baseline ES [/ha]`)) +
#   geom_point(aes(color = ES), size=0.5) + 
#   xlab("% Natural land use") +
#   geom_smooth(method = "gam") + facet_wrap(~ES, scales = "free", ncol = 4) + 
#   ggtitle("% natural area vs baseline ES") + egg::theme_article()

# lu.v.delta <- ggplot(dat_wider %>% filter(Landuse=="ntr_pland"),
#                      aes(Landuse_value, `Δ ES [/ha]`)) + 
#   geom_point(aes(color = ES), size=0.5) + 
#   xlab("% Natural land use") +
#   geom_smooth(method = "gam") + facet_wrap(~ES, scales = "free", ncol = 4) + 
#   ggtitle("% natural area vs. change in ES") + egg::theme_article()

# base.v.perc <- ggplot(dat_wider, aes(`Baseline ES [/ha]`, `% Δ ES`)) + 
#   geom_point(aes(color = ES), size=0.5) + geom_smooth(method = "gam") + 
#   facet_wrap(~ES, scales = "free", ncol = 4) +
#   ggtitle("ES Baseline vs ES % improvement") + 
#   egg::theme_article()

# base.v.permargin <- ggplot(dat_wider, aes(`Baseline ES [/ha]`, `Δ ES per ha margin`)) + 
#     geom_point(aes(color = ES), size=0.5) + geom_smooth(method = "gam") +
#   facet_wrap(~ES, scales = "free", ncol = 4) +
#     ggtitle("ES Baseline vs ES improvement per margin") + 
#     egg::theme_article()

# cowplot::plot_grid(lu.v.base, lu.v.delta, base.v.perc, base.v.permargin, ncol = 1)

```

```{r gams}
# fit some gams
dat.gam <- dat.long %>% filter(Landuse == "ntr_pland") %>% 
  group_by(response_type, ES) %>%
  nest() %>%
  # Fit GAM
  mutate(Mod = map(data, ~mgcv::gam(ES_value ~ s(Landuse_value), data = .x))) %>%
  # Get R2
  mutate(R2 = map_dbl(Mod, ~round(summary(.x)$r.sq, 2)))  

```

```{r natural}

ES_labels <- c(
  pollinators = "Pollinators \n[abundance/ha]",
  qb = "Water recharge \n[mm baseflow/ha]",
  sed_export = "Sediment export \n[metric tons/ha/year]",
  n_export = "Nutrient export \n[kg N/ha/year]"
)
response_type_labels <- c(
    base = "Baseline ES",
    delta = "\u0394 ES",
    pcmargin = "% \u0394 ES",
    delta.margin = "\u0394 ES per ha margin"
)

# paired plots of baseline and delta
ES_ntr <- ggplot(dat.long %>% 
         filter(Landuse == "ntr_pland" & response_type %in% c("base", "delta", "delta.margin")),
       aes(Landuse_value, ES_value)) + 
  geom_point(aes(color = ES), size = 0.5) + 
  geom_smooth(method="gam") +
  ggh4x::facet_grid2(response_type ~ ES,  scales = "free", independent = "y", 
                     switch = "y", 
                     labeller = labeller(ES = ES_labels,
                                         response_type = response_type_labels)) + 
  egg::theme_article() + 
  # Add label
  geom_label(data = dat.gam %>% 
         filter(response_type %in% c("base", "delta", "delta.margin")),
             aes(x = Inf, y = Inf,
                 label = paste("R2 = ", R2, sep = " ")),
             hjust = 1, vjust = 1) +
  xlab("Natural % landscape") + ylab("Ecosystem service (ES) response") +
  theme(strip.placement = "outside") +
  theme(legend.position = "none")

ggsave("Figs/ES_ntr.png", ES_ntr, bg="white", width = 7, height = 5, dpi=300)

```

```{r against_base}

dat.long.base <- dat.long %>% select(-ES_response) %>%
  pivot_wider(names_from = response_type, values_from = ES_value) %>%
  filter(Landuse == "ntr_pland") %>% select(-Landuse, -Landuse_value) %>%
  pivot_longer(delta:pcmargin, names_to = "change_type", values_to = "change_value") %>%
  mutate(change_type = factor(change_type,
                              levels = c("base", "delta", "pcmargin", "delta.margin")))

# fit some gams against baseline
dat.base.gam <- dat.long.base %>% 
  group_by(change_type, ES) %>%
  nest() %>%
  # Fit GAM
  mutate(Mod = map(data, ~mgcv::gam(change_value ~ s(base), data = .x))) %>%
  # Get R2
  mutate(R2 = map_dbl(Mod, ~round(summary(.x)$r.sq, 2)))  

# paired plots of baseline and delta
ES_base <- ggplot(dat.long.base %>% 
         filter(change_type %in% c("delta", "delta.margin")),
       aes(base, change_value)) + 
  geom_point(aes(color = ES), size = 0.5) + 
  geom_smooth(method="gam") +
  ggh4x::facet_grid2(change_type ~ ES,  scales = "free", independent = "y", 
                     switch = "both", 
                     labeller = labeller(ES = ES_labels,
                                         change_type = response_type_labels)) + 
  egg::theme_article() + 
  # Add label
  geom_label(data = dat.base.gam %>% 
         filter(change_type %in% c("delta", "delta.margin")),
             aes(x = Inf, y = Inf,
                 label = paste("R2 = ", R2, sep = " ")),
             hjust = 1, vjust = 1) +
  xlab("Baseline ecosystem service") + ylab("Ecosystem service change") +
  theme(strip.placement = "outside") +
  theme(legend.position = "none")

ggsave("Figs/ES_base.png", ES_base, bg="white", width = 7, height = 3.5, dpi=300)


```

## Patterns with annual agriculture

Looking at ES % change from baseline shows some stronger patterns than before for sediment and water recharge.

```{r compplots, fig.width = 8, fig.height=6, message =FALSE}

ggplot(dat.long %>% filter(Landuse == "ntr_pland"),
       aes(Landuse_value, ES_value)) + 
  geom_point(aes(color = ES), size = 0.5) + 
  geom_smooth(method="gam") +
  ggh4x::facet_grid2(response_type ~ ES,  scales = "free", independent = "y", 
                     switch = "y") + 
  egg::theme_article() + 
  # Add label
  geom_label(data = dat.gam,
             aes(x = Inf, y = Inf,
                 label = paste("R2 = ", R2, sep = " ")),
             hjust = 1, vjust = 1) +
  xlab("Natural % landscape") +
  ylab("ES value") + theme(strip.placement = "outside")
``` 

### Random forest model performance
*Pollinators:*

Baseline model

     r2 = `r round(pollinators_base.rf$r.squared, 3)`
     
     OOB MSE = `r format(round(pollinators_base.rf$prediction.error, 5),scientific=F)`
     
Delta model

     r2 = `r round(pollinators_delta.rf$r.squared, 3)`
     
     OOB MSE = `r format(round(pollinators_delta.rf$prediction.error, 5),scientific=F)`

Percent change model

     r2 = `r round(pollinators_pc.rf$r.squared, 3)`
     
     OOB MSE = `r format(round(pollinators_pc.rf$prediction.error, 5),scientific=F)`
     
Per margin model

     r2 = `r round(pollinators_margin.rf$r.squared, 3)`
     
     OOB MSE = `r format(round(pollinators_margin.rf$prediction.error, 5),scientific=F)`

*Nitrogen export:*

Baseline model

     r2 = `r round(n_export_base.rf$r.squared, 3)`
     
     OOB MSE = `r format(round(n_export_base.rf$prediction.error, 5),scientific=F)`

Delta model

     r2 = `r round(n_export_delta.rf$r.squared, 3)`
     
     OOB MSE = `r format(round(n_export_delta.rf$prediction.error, 5),scientific=F)`

Percent change model

     r2 = `r round(n_export_pc.rf$r.squared, 3)`
     
     OOB MSE = `r format(round(n_export_pc.rf$prediction.error, 5),scientific=F)`
     
Per margin model

     r2 = `r round(n_export_margin.rf$r.squared, 3)`
     
     OOB MSE = `r format(round(n_export_margin.rf$prediction.error, 5),scientific=F)`

*Sediment export:*

Baseline model

     r2 = `r round(sed_export_base.rf$r.squared, 3)`
     
     OOB MSE = `r format(round(sed_export_base.rf$prediction.error, 5),scientific=F)`

Delta model

     r2 = `r round(sed_export_delta.rf$r.squared, 3)`
     
     OOB MSE = `r format(round(sed_export_delta.rf$prediction.error, 5),scientific=F)`

Percent change model

     r2 = `r round(sed_export_pc.rf$r.squared, 3)`
     
     OOB MSE = `r format(round(sed_export_pc.rf$prediction.error, 5),scientific=F)`
     
Per margin model

     r2 = `r round(sed_export_margin.rf$r.squared, 3)`
     
     OOB MSE = `r format(round(sed_export_margin.rf$prediction.error, 5),scientific=F)`

*Baseflow:*

Baseline model

     r2 = `r round(qb_base.rf$r.squared, 3)`
     
     OOB MSE = `r format(round(qb_base.rf$prediction.error, 5),scientific=F)`

Delta model

     r2 = `r round(qb_delta.rf$r.squared, 3)`
     
     OOB MSE = `r format(round(qb_delta.rf$prediction.error, 5),scientific=F)`

Percent change model

     r2 = `r round(qb_pc.rf$r.squared, 3)`
     
     OOB MSE = `r format(round(qb_pc.rf$prediction.error, 5),scientific=F)`
     
Per margin model

     r2 = `r round(qb_margin.rf$r.squared, 3)`
     
     OOB MSE = `r format(round(qb_margin.rf$prediction.error, 5),scientific=F)`
     
### Variable importance

```{r importance, fig.height=16, fig.width=12, include=FALSE}
library(ggpattern)

metric.descr <- read.csv("data/lsm_description.csv")

# collect data
all.impdat <- list(
  list(pollinators_base.p, n_export_base.p, sed_export_base.p, qb_base.p),
  list(pollinators_delta.p, n_export_delta.p, sed_export_delta.p, qb_delta.p),
  list(pollinators_pc.p, n_export_pc.p, sed_export_pc.p, qb_pc.p),
  list(pollinators_margin.p, n_export_margin.p, sed_export_margin.p, qb_margin.p)
) %>% lapply(\(x) setNames(x, c("Pollinators", "N export", "Sed export", "Baseflow"))) %>%
  lapply(\(x) lapply(x,\(y)rownames_to_column(y, "metric"))) %>%  # wow this is ridiculous
  lapply(\(x)bind_rows(x,.id="ES")) %>%
  setNames(nm = c("baseline", "delta", "percent_change", "per_margin")) %>%
  bind_rows(.id="model")

imp.df <- all.impdat %>%
  mutate(sig = ifelse(pvalue <0.05, TRUE, FALSE)) %>%
  mutate(metric2 = tidytext::reorder_within(metric, importance,
                                           within = list(ES, model))) %>%
  left_join(metric.descr %>% select(lsm, type), by = c("metric"="lsm")) %>%
  mutate(type = case_match(
    metric,
    "field_pland" ~ "area and edge metric",
    "field_frac_mn" ~ "shape metric",
    "field_frac_cv" ~ "shape metric",
    "prn_pagr" ~ "area and edge metric",
    .default = type
  ))

ggplot(imp.df, aes(metric2, importance, fill = type, pattern = sig)) +
  geom_bar_pattern(stat="identity",
                   color = "black",
                   pattern_fill = "black",
                   pattern_angle = 45,
                   pattern_density = 0.1,
                   pattern_spacing = 0.025,
                   pattern_key_scale_factor = 0.6) +
  tidytext::scale_x_reordered() +
  coord_flip() +
  ggh4x::facet_grid2(ES~model, scales = "free", independent = "all") +
  scale_pattern_manual(values = c("TRUE" = "none", "FALSE" = "stripe")) +
  egg::theme_article() +
  theme(legend.position = "top") +
  guides(pattern = guide_legend(override.aes = list(fill = "white")),
         fill = guide_legend(override.aes = list(pattern = "none")))


```

```{r importance_shap}
load("data/ES-RF-all-shapimportance.RDA")   # prediction objects from iml package
load("data/ES-RF-delta-shapimportance.RDA")
library(shapviz)
library(gridExtra)

```

#### Pollinators

```{r pollinator_shaps, fig.width=12, fig.height=3}

svb.pol.base <- sv_importance(ex.polbase.glb, kind = "bee", max_display = 5) +
  egg::theme_article() + theme(legend.position = "none")
svb.pol.delta <- sv_importance(ex.poldelta.glb, kind = "bee", max_display = 5) + 
  egg::theme_article() + theme(legend.position = "none")
svb.pol.margin <- sv_importance(ex.polmarg.glb, kind = "bee", max_display = 5) +
  egg::theme_article() + theme(legend.position = "none")

cowplot::plot_grid(
  svb.pol.base, svb.pol.delta,svb.pol.margin,
  nrow = 1,
  labels = c("(a)", "(b)", "(c)")
)

```

#### Nutrient export

```{r n_export_shaps, fig.width=12, fig.height=3}

svb.nexp.base <- sv_importance(ex.nexpbase.glb, kind = "bee", max_display = 5) +
  egg::theme_article() + theme(legend.position = "none")
svb.nexp.delta <- sv_importance(ex.nexpdelta.glb, kind = "bee", max_display = 5) + 
  egg::theme_article() + theme(legend.position = "none")
svb.nexp.margin <- sv_importance(ex.nexpmarg.glb, kind = "bee", max_display = 5) +
  egg::theme_article() + theme(legend.position = "none")

cowplot::plot_grid(
  svb.nexp.base, svb.nexp.delta,svb.nexp.margin,
  nrow = 1,
  labels = c("(a)", "(b)", "(c)")
)

```

#### Sediment export

```{r sed_exp_shaps, fig.width=12, fig.height=3}

svb.sed.base <- sv_importance(ex.sedbase.glb, kind = "bee", max_display = 5) +
  egg::theme_article() + theme(legend.position = "none")
svb.sed.delta <- sv_importance(ex.seddelta.glb, kind = "bee", max_display = 5) + 
  egg::theme_article() + theme(legend.position = "none")
svb.sed.margin <- sv_importance(ex.sedmarg.glb, kind = "bee", max_display = 5) +
  egg::theme_article() + theme(legend.position = "none")

cowplot::plot_grid(
  svb.sed.base, svb.sed.delta,svb.sed.margin,
  nrow = 1,
  labels = c("(a)", "(b)", "(c)")
)

```

#### Baseflow

```{r qb_shaps, fig.width=12, fig.height=3}

svb.qb.base <- sv_importance(ex.qbbase.glb, kind = "bee", max_display = 5) +
  egg::theme_article() + theme(legend.position = "none")
svb.qb.delta <- sv_importance(ex.qbdelta.glb, kind = "bee", max_display = 5) + 
  egg::theme_article() + theme(legend.position = "none")
svb.qb.margin <- sv_importance(ex.qbmarg.glb, kind = "bee", max_display = 5) +
  egg::theme_article() + theme(legend.position = "none")

cowplot::plot_grid(
  svb.qb.base, svb.qb.delta,svb.qb.margin,
  nrow = 1,
  labels = c("(a)", "(b)", "(c)")
)

```

### Across ecosystem services

```{r shap_fun}

# function to process the shap data

sv_fun <- function(shapviz_list, metric){
  svd.df <- lapply(shapviz_list, \(x) shapviz::sv_dependence(x, v=metric)$data) %>%
    setNames(names(shapviz_list)) %>% bind_rows(.id = "ES")
}

# base data
base.glb <- list(pollinators = ex.polbase.glb, qb = ex.qbbase.glb, 
                 sed_export = ex.sedbase.glb, n_export = ex.nexpbase.glb)

# delta data
delta.glb <- list(pollinators = ex.poldelta.glb, qb = ex.qbdelta.glb, 
                 sed_export = ex.seddelta.glb, n_export = ex.nexpdelta.glb)

# delta data
margin.glb <- list(pollinators = ex.polmarg.glb, qb = ex.qbmarg.glb, 
                 sed_export = ex.sedmarg.glb, n_export = ex.nexpmarg.glb)

```

#### Baseline: Across percent natural landscape

```{r across_ntr, fig.height=3, fig.width=9}

shapval_ntr.base <- sv_fun(base.glb, "ntr_pland")
shapval_ntr.delta <- sv_fun(delta.glb, "ntr_pland")
shapval_ntr.marg <- sv_fun(margin.glb, "ntr_pland")

shapval_ntr.all <- list(base = shapval_ntr.base, delta = shapval_ntr.delta, margin = shapval_ntr.marg) %>%
  bind_rows(.id = "response")

ggplot(shapval_ntr.all, aes(ntr_pland, shap, color = ES)) + geom_point(size=0.5) + 
  geom_smooth(method = "gam", se = FALSE) + egg::theme_article() +
  facet_wrap(~response, scales = "free") + theme(legend.position = "bottom")

```

#### Baseline: Across annual agriculture edge density

Important for sediment export (and pollinator delta)

```{r across_annagr_ed, fig.height=3, fig.width=9}

shapval_anned.base <- sv_fun(base.glb, "ann_agr_ed")
shapval_anned.delta <- sv_fun(delta.glb, "ann_agr_ed")
shapval_anned.marg <- sv_fun(margin.glb, "ann_agr_ed")

shapval_anned.all <- list(base = shapval_anned.base, 
                          delta = shapval_anned.delta, 
                          margin = shapval_anned.marg) %>%
  bind_rows(.id = "response")

ggplot(shapval_anned.all, aes(ann_agr_ed, shap, color = ES)) + geom_point(size=0.5) + 
  geom_smooth(method = "gam", se = FALSE) + egg::theme_article() +
  facet_wrap(~response, scales = "free") + theme(legend.position = "bottom")

```

#### Baseline: Across annual agriculture cohesion

Important for baseflow

```{r across_annagr_cohesion, fig.height=3, fig.width=9}

shapval_anncoh.base <- sv_fun(base.glb, "ann_agr_cohesion")
shapval_anncoh.delta <- sv_fun(delta.glb, "ann_agr_cohesion")
shapval_anncoh.marg <- sv_fun(margin.glb, "ann_agr_cohesion")

shapval_anncoh.all <- list(base = shapval_anncoh.base, 
                          delta = shapval_anncoh.delta, 
                          margin = shapval_anncoh.marg) %>%
  bind_rows(.id = "response")

ggplot(shapval_anncoh.all, aes(ann_agr_cohesion, shap, color = ES)) + geom_point(size=0.5) + 
  geom_smooth(method = "gam", se = FALSE) + egg::theme_article() +
  facet_wrap(~response, scales = "free") + theme(legend.position = "bottom")

```

#### Baseline: Across natural land cover edge density

Important for nutrient export

```{r across_ntr_ed, fig.height=3, fig.width=9}

shapval_ntred.base <- sv_fun(base.glb, "ntr_ed")
shapval_ntred.delta <- sv_fun(delta.glb, "ntr_ed")
shapval_ntred.marg <- sv_fun(margin.glb, "ntr_ed")

shapval_ntred.all <- list(base = shapval_ntred.base, 
                          delta = shapval_ntred.delta, 
                          margin = shapval_ntred.marg) %>%
  bind_rows(.id = "response")

ggplot(shapval_ntred.all, aes(ntr_ed, shap, color = ES)) + geom_point(size=0.5) + 
  geom_smooth(method = "gam", se = FALSE) + egg::theme_article() +
  facet_wrap(~response, scales = "free") + theme(legend.position = "bottom")

```