---
title: "Percent change"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r load}
library(tidyverse)
library(iml)
library(ranger)
load("data/ES-RF-all.RDA")
load("data/ES-RF-delta.RDA")
```

```{r metric_cor}

renamefun <- function(x){
  x %>% dplyr::rename(
  annual_perc_land = "ann_agr_pland",
  field_perc_land = "field_pland",
  nature_perc_land = "ntr_pland",
  perenn_perc_land = "prn_agr_pland",
  landuse_interjuxt = "aglu_iji",
  landuse_contagion = "aglu_contag",
  annual_cohesion = "ann_agr_cohesion",
  nature_cohesion = "ntr_cohesion",
  perenn_cohesion = "prn_agr_cohesion",
  landuse_edge_dens = "aglu_ed",
  landuse_frac_dim = "aglu_frac_mn",
  annual_patch_area = "ann_agr_area_mn",
  annual_edge_dens = "ann_agr_ed",
  annual_frac_dim = "ann_agr_frac_mn",
  annual_patch_dens = "ann_agr_pd",
  nature_patch_area = "ntr_area_mn",
  nature_edge_dens = "ntr_ed",
  nature_frac_dim = "ntr_frac_mn",
  nature_patch_dens = "ntr_pd",
  perenn_perc_agri = "prn_pagr",
  perenn_patch_area = "prn_agr_area_mn",
  perenn_edge_dens = "prn_agr_ed",
  perenn_frac_dim = "prn_agr_frac_mn",
  perenn_patch_dens = "prn_agr_pd",
  field_patch_area = "field_area_mn",
  field_edge_dens = "field_ed",
  field_frac_dim = "field_frac_mn",
  field_patch_dens = "field_pd",
  landuse_shan_div = "lu_shdi",
  crop_shan_div = "crop_shdi"
)
}

dat.rename <- renamefun(dat)

cor.met <- corrplot::corrplot(cor(dat.rename[,17:42]),
                   order = "hclust", hclust.method = "complete",addrect = 2,
                   # type = "upper", 
                   tl.col = 'black', tl.srt = 55, tl.cex = 0.8, tl.pos = "lt",
                   number.cex = 0.5, number.digits=1)

# cor.met$arg$type <- "full"

# corrplot::corrRect.hclust(cor.met, k = 2, method = "complete")

cor.met2 <- {corrplot::corrplot.mixed(cor(dat.rename[,17:42]),
                   order = "hclust", hclust.method = "complete",addrect = 2,
                   # type = "upper", 
                   tl.col = 'black', tl.srt = 55, tl.cex = 0.8, tl.pos = "lt",
                   number.cex = 0.6, number.digits=2, number.col="black");
  recordPlot()
  }

# ggsave("Figs/SupFig_corplot.png", replayPlot(cor.met2), height = 10, width = 10, units = "in", bg = "white", dpi = 300)

```

## Ecosystem service improvements depend on the baseline condition

```{r data_setup}


dat.long <- dat %>% 
  pivot_longer(cols = c(1:16), names_to = "ES_response", values_to = "ES_value") %>%
  pivot_longer(cols = c("ntr_pland", "ann_agr_pland"), names_to = "Landuse",
               values_to = "Landuse_value") %>%
  select(ES_response, ES_value, Landuse, Landuse_value) %>%
  mutate(ES = gsub("delta_|_base|_(incr|red).margin|pc_|_max.margin","", ES_response)) %>%
  mutate(ES = case_match(ES,
                         "sed" ~ "sed_export",
                         "pol" ~ "pollinators",
                         "nexp" ~ "n_export",
                         .default = ES)) %>%
  mutate(response_type =
           gsub("sed|export|n_|pollinators|qb|_|pol|nexp|red.|incr.|max","",ES_response)) %>%
  # mutate(response_type = case_match(
  #   response_type,
  #   "base" ~ "Baseline ES [/ha]",
  #   "delta" ~ "\u0394 ES [/ha]",
  #   "pcmargin" ~ "% \u0394 ES",
  #   "delta.margin" ~ "\u0394 ES per ha margin"
  # )) %>%
  mutate(response_type = factor(response_type, 
                                levels = c("base",
                                           "delta",
                                           "pcmargin",
                                           "delta.margin")),
         ES = factor(ES, levels = c("pollinators","qb", "sed_export", "n_export")))

```

Ecosystem service improvements depend on the baseline condition, except for pollination.

This pattern holds even if we factor out margin area. 

The trend is different if we look at % change in ecosystem services from baseline, per margin area

```{r percentchangedemo, message=FALSE, fig.width = 8, fig.height=9}

# dat_wider <- dat.long %>% select(-ES_response) %>% 
#   pivot_wider(names_from = response_type, values_from = ES_value)

# lu.v.base <- ggplot(dat_wider %>% filter(Landuse=="ntr_pland"), 
#                        aes(Landuse_value,`Baseline ES [/ha]`)) +
#   geom_point(aes(color = ES), size=0.5) + 
#   xlab("% Natural land use") +
#   geom_smooth(method = "gam") + facet_wrap(~ES, scales = "free", ncol = 4) + 
#   ggtitle("% natural area vs baseline ES") + egg::theme_article()

# lu.v.delta <- ggplot(dat_wider %>% filter(Landuse=="ntr_pland"),
#                      aes(Landuse_value, `Δ ES [/ha]`)) + 
#   geom_point(aes(color = ES), size=0.5) + 
#   xlab("% Natural land use") +
#   geom_smooth(method = "gam") + facet_wrap(~ES, scales = "free", ncol = 4) + 
#   ggtitle("% natural area vs. change in ES") + egg::theme_article()

# base.v.perc <- ggplot(dat_wider, aes(`Baseline ES [/ha]`, `% Δ ES`)) + 
#   geom_point(aes(color = ES), size=0.5) + geom_smooth(method = "gam") + 
#   facet_wrap(~ES, scales = "free", ncol = 4) +
#   ggtitle("ES Baseline vs ES % improvement") + 
#   egg::theme_article()

# base.v.permargin <- ggplot(dat_wider, aes(`Baseline ES [/ha]`, `Δ ES per ha margin`)) + 
#     geom_point(aes(color = ES), size=0.5) + geom_smooth(method = "gam") +
#   facet_wrap(~ES, scales = "free", ncol = 4) +
#     ggtitle("ES Baseline vs ES improvement per margin") + 
#     egg::theme_article()

# cowplot::plot_grid(lu.v.base, lu.v.delta, base.v.perc, base.v.permargin, ncol = 1)

```

```{r gams}

# function to get gam fit info
k.check.fun <- function(b, k.sample = 5000, k.rep = 200) {
  mgcv:::k.check(b, subsample = k.sample, n.rep = k.rep)
}

mae.fun <- function(gmod){
  mean(abs(residuals(gmod)))
}
mse.fun <- function(gmod){
  mean((residuals(gmod))^2)
}

# fit some gams
dat.gam <- dat.long %>% filter(Landuse == "ntr_pland") %>% 
  group_by(response_type, ES) %>%
  nest() %>%
  # Fit GAM
  mutate(Mod = map(data, ~mgcv::gam(ES_value ~ s(Landuse_value), data = .x))) %>%
  # Get R2
  mutate(R2 = map_dbl(Mod, ~round(summary(.x)$r.sq, 2)),
         mae = map_dbl(Mod, ~round(mae.fun(.x), 3)),
         mse = map_dbl(Mod, ~round(mse.fun(.x), 3))
         ) %>%
  mutate(
    kp =  map_dbl(Mod, ~round(k.check.fun(.x)[1], 3)),
    edf =  map_dbl(Mod, ~round(k.check.fun(.x)[2], 3)),
    kindex =  map_dbl(Mod, ~round(k.check.fun(.x)[3], 3)),
    pvalue =  map_dbl(Mod, ~round(k.check.fun(.x)[4], 3))
  )

# write.csv(dat.gam %>% select(ES, response_type, R2, mae, mse, kp, edf, kindex, pvalue), "gam_ntr2.csv", row.names = FALSE)

```

```{r natural}

ES_labels <- c(
  pollinators = "Wild bee abundance \n[abundance/ha]",
  qb = "Groundwater recharge \n[mm baseflow/ha]",
  sed_export = "Sediment runoff \n[metric tons/ha/year]",
  n_export = "Nutrient runoff \n[kg N/ha/year]"
)
response_type_labels <- c(
    base = "Baseline ES",
    delta = "\u0394 ES (landscape)",
    pcmargin = "% \u0394 ES",
    delta.margin = "\u0394 ES (margin)"
)

# paired plots of baseline and delta
ES_ntr <- ggplot(dat.long %>% 
         filter(Landuse == "ntr_pland" & response_type %in% c("base", "delta", "delta.margin")),
       aes(Landuse_value, ES_value)) + 
  geom_point(aes(color = ES), size = 0.5) + scale_color_viridis_d(option = "turbo") +
  geom_smooth(method="gam") +
  ggh4x::facet_grid2(response_type ~ ES,  scales = "free", independent = "y", 
                     switch = "y", 
                     labeller = labeller(ES = ES_labels,
                                         response_type = response_type_labels)) + 
  egg::theme_article() + 
  # Add label
  # geom_label(data = dat.gam %>% 
  #        filter(response_type %in% c("base", "delta", "delta.margin")),
  #            aes(x = Inf, y = Inf,
  #                label = paste("R2 = ", R2, sep = " ")),
  #            hjust = 1, vjust = 1) +
  xlab("% Natural area") + ylab("Ecosystem service (ES) response") +
  theme(strip.placement = "outside") +
  theme(legend.position = "none")

ES_ntr_tag <- ES_ntr %>% egg::tag_facet() + 
  theme(strip.text = element_text(size = 8, margin = margin(r=1,b=6,l=6)))

# ggsave("Figs/ES_ntr.png", ES_ntr_tag, bg="white", width = 8, height = 5, dpi=300)

```

```{r against_base}

ES_labels_flip <- c(
  pollinators = "[abundance/ha]\nPollinators",
  qb = "[mm baseflow/ha]\nGroundwater recharge",
  sed_export = "[metric tons/ha/year]\nSediment runoff",
  n_export = "[kg N/ha/year]\nNutrient runoff"
)

dat.long.base <- dat.long %>% select(-ES_response) %>%
  pivot_wider(names_from = response_type, values_from = ES_value) %>%
  filter(Landuse == "ntr_pland") %>% select(-Landuse, -Landuse_value) %>%
  pivot_longer(delta:pcmargin, names_to = "change_type", values_to = "change_value") %>%
  mutate(change_type = factor(change_type,
                              levels = c("base", "delta", "pcmargin", "delta.margin")))

# fit some gams against baseline
dat.base.gam <- dat.long.base %>% 
  group_by(change_type, ES) %>%
  nest() %>%
  # Fit GAM
  mutate(Mod = map(data, ~mgcv::gam(change_value ~ s(base), data = .x))) %>%
  # Get R2
  mutate(R2 = map_dbl(Mod, ~round(summary(.x)$r.sq, 2)),
         mae = map_dbl(Mod, ~round(mae.fun(.x), 3))) %>%
  mutate(
    kp =  map_dbl(Mod, ~round(k.check.fun(.x)[1], 3)),
    edf =  map_dbl(Mod, ~round(k.check.fun(.x)[2], 3)),
    kindex =  map_dbl(Mod, ~round(k.check.fun(.x)[3], 3)),
    pvalue =  map_dbl(Mod, ~round(k.check.fun(.x)[4], 3))
  )

# paired plots of baseline and delta
ES_base <- ggplot(dat.long.base %>% 
         filter(change_type %in% c("delta", "delta.margin")),
       aes(base, change_value)) + 
  geom_point(aes(color = ES), size = 0.5) + scale_color_viridis_d(option = "turbo") +
  geom_smooth(method="gam") +
  ggh4x::facet_grid2(change_type ~ ES,  scales = "free", independent = "y", 
                     switch = "both", 
                     labeller = labeller(ES = ES_labels_flip,
                                         change_type = response_type_labels)) + 
  egg::theme_article() + 
  # Add label
  geom_label(data = dat.base.gam %>% 
         filter(change_type %in% c("delta", "delta.margin")),
             aes(x = Inf, y = Inf,
                 label = paste("R2 = ",sprintf("%.2f", R2), sep = "")),
             hjust = 1, vjust = 1) +
  xlab("Baseline ecosystem service") + ylab("Ecosystem service (ES) improvement\nin maximum scenario") +
  theme(strip.placement = "outside") +
  theme(legend.position = "none")

ES_base_tag <- ES_base %>% egg::tag_facet() + 
  theme(strip.text = element_text(size = 8, margin = margin(r=1,b=6,l=6)))

ES_base_tag

# ggsave("Figs/ES_base.png", ES_base_tag, bg="white", width = 7, height = 3.5, dpi=300)

# write.csv(dat.base.gam %>% select(ES, change_type, R2, mae, kp, edf, kindex, pvalue), "gam_base.csv", row.names = FALSE)

```

## Patterns with annual agriculture

Looking at ES % change from baseline shows some stronger patterns than before for sediment and water recharge.

```{r compplots, fig.width = 8, fig.height=6, include =FALSE}

ggplot(dat.long %>% filter(Landuse == "ntr_pland"),
       aes(Landuse_value, ES_value)) + 
  geom_point(aes(color = ES), size = 0.5) + 
  geom_smooth(method="gam") +
  ggh4x::facet_grid2(response_type ~ ES,  scales = "free", independent = "y", 
                     switch = "y") + 
  egg::theme_article() + 
  # Add label
  geom_label(data = dat.gam,
             aes(x = Inf, y = Inf,
                 label = paste("R2 = ", R2, sep = " ")),
             hjust = 1, vjust = 1) +
  xlab("Natural % landscape") +
  ylab("ES value") + theme(strip.placement = "outside")
``` 

### Random forest model performance
*Pollinators:*

Baseline model

     r2 = `r round(pollinators_base.rf$r.squared, 3)`
     
     OOB MSE = `r format(round(pollinators_base.rf$prediction.error, 5),scientific=F)`
     
Delta model

     r2 = `r round(pollinators_delta.rf$r.squared, 3)`
     
     OOB MSE = `r format(round(pollinators_delta.rf$prediction.error, 5),scientific=F)`

Percent change model

     r2 = `r round(pollinators_pc.rf$r.squared, 3)`
     
     OOB MSE = `r format(round(pollinators_pc.rf$prediction.error, 5),scientific=F)`
     
Per margin model

     r2 = `r round(pollinators_margin.rf$r.squared, 3)`
     
     OOB MSE = `r format(round(pollinators_margin.rf$prediction.error, 5),scientific=F)`

*Nitrogen export:*

Baseline model

     r2 = `r round(n_export_base.rf$r.squared, 3)`
     
     OOB MSE = `r format(round(n_export_base.rf$prediction.error, 5),scientific=F)`

Delta model

     r2 = `r round(n_export_delta.rf$r.squared, 3)`
     
     OOB MSE = `r format(round(n_export_delta.rf$prediction.error, 5),scientific=F)`

Percent change model

     r2 = `r round(n_export_pc.rf$r.squared, 3)`
     
     OOB MSE = `r format(round(n_export_pc.rf$prediction.error, 5),scientific=F)`
     
Per margin model

     r2 = `r round(n_export_margin.rf$r.squared, 3)`
     
     OOB MSE = `r format(round(n_export_margin.rf$prediction.error, 5),scientific=F)`

*Sediment export:*

Baseline model

     r2 = `r round(sed_export_base.rf$r.squared, 3)`
     
     OOB MSE = `r format(round(sed_export_base.rf$prediction.error, 5),scientific=F)`

Delta model

     r2 = `r round(sed_export_delta.rf$r.squared, 3)`
     
     OOB MSE = `r format(round(sed_export_delta.rf$prediction.error, 5),scientific=F)`

Percent change model

     r2 = `r round(sed_export_pc.rf$r.squared, 3)`
     
     OOB MSE = `r format(round(sed_export_pc.rf$prediction.error, 5),scientific=F)`
     
Per margin model

     r2 = `r round(sed_export_margin.rf$r.squared, 3)`
     
     OOB MSE = `r format(round(sed_export_margin.rf$prediction.error, 5),scientific=F)`

*Baseflow:*

Baseline model

     r2 = `r round(qb_base.rf$r.squared, 3)`
     
     OOB MSE = `r format(round(qb_base.rf$prediction.error, 5),scientific=F)`

Delta model

     r2 = `r round(qb_delta.rf$r.squared, 3)`
     
     OOB MSE = `r format(round(qb_delta.rf$prediction.error, 5),scientific=F)`

Percent change model

     r2 = `r round(qb_pc.rf$r.squared, 3)`
     
     OOB MSE = `r format(round(qb_pc.rf$prediction.error, 5),scientific=F)`
     
Per margin model

     r2 = `r round(qb_margin.rf$r.squared, 3)`
     
     OOB MSE = `r format(round(qb_margin.rf$prediction.error, 5),scientific=F)`
     
### Variable importance

```{r importance, fig.height=16, fig.width=12, include=FALSE}
library(ggpattern)

metric.descr <- read.csv("data/lsm_description.csv")

# collect data
all.impdat <- list(
  list(pollinators_base.p, n_export_base.p, sed_export_base.p, qb_base.p),
  list(pollinators_delta.p, n_export_delta.p, sed_export_delta.p, qb_delta.p),
  list(pollinators_pc.p, n_export_pc.p, sed_export_pc.p, qb_pc.p),
  list(pollinators_margin.p, n_export_margin.p, sed_export_margin.p, qb_margin.p)
) %>% lapply(\(x) setNames(x, c("Pollinators", "N export", "Sed export", "Baseflow"))) %>%
  lapply(\(x) lapply(x,\(y)rownames_to_column(y, "metric"))) %>%  # wow this is ridiculous
  lapply(\(x)bind_rows(x,.id="ES")) %>%
  setNames(nm = c("baseline", "delta", "percent_change", "per_margin")) %>%
  bind_rows(.id="model")

imp.df <- all.impdat %>%
  mutate(sig = ifelse(pvalue <0.05, TRUE, FALSE)) %>%
  mutate(metric2 = tidytext::reorder_within(metric, importance,
                                           within = list(ES, model))) %>%
  left_join(metric.descr %>% select(lsm, type), by = c("metric"="lsm")) %>%
  mutate(type = case_match(
    metric,
    "field_pland" ~ "area and edge metric",
    "field_frac_mn" ~ "shape metric",
    "field_frac_cv" ~ "shape metric",
    "prn_pagr" ~ "area and edge metric",
    .default = type
  ))

ggplot(imp.df, aes(metric2, importance, fill = type, pattern = sig)) +
  geom_bar_pattern(stat="identity",
                   color = "black",
                   pattern_fill = "black",
                   pattern_angle = 45,
                   pattern_density = 0.1,
                   pattern_spacing = 0.025,
                   pattern_key_scale_factor = 0.6) +
  tidytext::scale_x_reordered() +
  coord_flip() +
  ggh4x::facet_grid2(ES~model, scales = "free", independent = "all") +
  scale_pattern_manual(values = c("TRUE" = "none", "FALSE" = "stripe")) +
  egg::theme_article() +
  theme(legend.position = "top") +
  guides(pattern = guide_legend(override.aes = list(fill = "white")),
         fill = guide_legend(override.aes = list(pattern = "none")))


```

```{r importance_shap}
load("data/ES-RF-all-shapimportance.RDA")   # prediction objects from iml package
load("data/ES-RF-delta-shapimportance.RDA")
library(shapviz)
library(gridExtra)

```

```{r rename_factors_fun}

rename_fac_fun <- function(x){
  x$data <- x$data %>% 
    dplyr::mutate(
      Var2 = forcats::fct_recode(
        Var2,
        annual_perc_land = "ann_agr_pland",
        field_perc_land = "field_pland",
        nature_perc_land = "ntr_pland",
        perenn_perc_land = "prn_agr_pland",
        landuse_interjuxt = "aglu_iji",
        landuse_contagion = "aglu_contag",
        annual_cohesion = "ann_agr_cohesion",
        nature_cohesion = "ntr_cohesion",
        perenn_cohesion = "prn_agr_cohesion",
        landuse_edge_dens = "aglu_ed",
        landuse_frac_dim = "aglu_frac_mn",
        annual_patch_area = "ann_agr_area_mn",
        annual_edge_den = "ann_agr_ed",
        annual_frac_dim = "ann_agr_frac_mn",
        annual_patch_dens = "ann_agr_pd",
        nature_patch_area = "ntr_area_mn",
        nature_edge_dens = "ntr_ed",
        nature_frac_dim = "ntr_frac_mn",
        nature_patch_dens = "ntr_pd",
        perenn_perc_agri = "prn_pagr",
        perenn_patch_area = "prn_agr_area_mn",
        perenn_edge_dens = "prn_agr_ed",
        perenn_frac_dim = "prn_agr_frac_mn",
        perenn_patch_dens = "prn_agr_pd",
        field_patch_area = "field_area_mn",
        field_edge_dens = "field_ed",
        field_frac_dim = "field_frac_mn",
        field_patch_dens = "field_pd",
        landuse_shan_div = "lu_shdi",
        crop_shan_div = "crop_shdi"
        )
      ) %>% 
    dplyr::mutate(
      feature = forcats::fct_recode(
        feature,
        annual_perc_land = "ann_agr_pland",
        field_perc_land = "field_pland",
        nature_perc_land = "ntr_pland",
        perenn_perc_land = "prn_agr_pland",
        landuse_interjuxt = "aglu_iji",
        landuse_contagion = "aglu_contag",
        annual_cohesion = "ann_agr_cohesion",
        nature_cohesion = "ntr_cohesion",
        perenn_cohesion = "prn_agr_cohesion",
        landuse_edge_dens = "aglu_ed",
        landuse_frac_dim = "aglu_frac_mn",
        annual_patch_area = "ann_agr_area_mn",
        annual_edge_den = "ann_agr_ed",
        annual_frac_dim = "ann_agr_frac_mn",
        annual_patch_dens = "ann_agr_pd",
        nature_patch_area = "ntr_area_mn",
        nature_edge_dens = "ntr_ed",
        nature_frac_dim = "ntr_frac_mn",
        nature_patch_dens = "ntr_pd",
        perenn_perc_agri = "prn_pagr",
        perenn_patch_area = "prn_agr_area_mn",
        perenn_edge_dens = "prn_agr_ed",
        perenn_frac_dim = "prn_agr_frac_mn",
        perenn_patch_dens = "prn_agr_pd",
        field_patch_area = "field_area_mn",
        field_edge_dens = "field_ed",
        field_frac_dim = "field_frac_mn",
        field_patch_dens = "field_pd",
        landuse_shan_div = "lu_shdi",
        crop_shan_div = "crop_shdi"
        )
      )
    x$layers[[3]]$data <- x$layers[[3]]$data %>% 
    dplyr::mutate(
      feature = forcats::fct_recode(
        feature,
        annual_perc_land = "ann_agr_pland",
        field_perc_land = "field_pland",
        nature_perc_land = "ntr_pland",
        perenn_perc_land = "prn_agr_pland",
        landuse_interjuxt = "aglu_iji",
        landuse_contagion = "aglu_contag",
        annual_cohesion = "ann_agr_cohesion",
        nature_cohesion = "ntr_cohesion",
        perenn_cohesion = "prn_agr_cohesion",
        landuse_edge_dens = "aglu_ed",
        landuse_frac_dim = "aglu_frac_mn",
        annual_patch_area = "ann_agr_area_mn",
        annual_edge_den = "ann_agr_ed",
        annual_frac_dim = "ann_agr_frac_mn",
        annual_patch_dens = "ann_agr_pd",
        nature_patch_area = "ntr_area_mn",
        nature_edge_dens = "ntr_ed",
        nature_frac_dim = "ntr_frac_mn",
        nature_patch_dens = "ntr_pd",
        perenn_perc_agri = "prn_pagr",
        perenn_patch_area = "prn_agr_area_mn",
        perenn_edge_dens = "prn_agr_ed",
        perenn_frac_dim = "prn_agr_frac_mn",
        perenn_patch_dens = "prn_agr_pd",
        field_patch_area = "field_area_mn",
        field_edge_dens = "field_ed",
        field_frac_dim = "field_frac_mn",
        field_patch_dens = "field_pd",
        landuse_shan_div = "lu_shdi",
        crop_shan_div = "crop_shdi"
        )
      )
  x
}


```

#### Pollinators

```{r pollinator_shaps, fig.width=12, fig.height=7}

ex.polbase.glb.rename <- ex.polbase.glb
ex.poldelta.glb.rename <- ex.poldelta.glb
ex.polmarg.glb.rename <- ex.polmarg.glb

svb.pol.base <- sv_importance(ex.polbase.glb, kind = "bee", max_display = 30,
                              viridis_args = list(option = "viridis"), size = 0.5,
                              show_numbers = TRUE, number_size = 2.7, format_fun = \(x) format(round(x,3),nsmall=3,)) +
  egg::theme_article() + theme(legend.position = "none") + 
  xlab("Contribution to baseline wild bee abundance")
svb.pol.delta <- sv_importance(ex.poldelta.glb, kind = "bee", max_display = 30,
                              viridis_args = list(option = "viridis"), size = 0.5,
                              show_numbers = TRUE, number_size = 2.7, format_fun = \(x) format(round(x,3),nsmall=3,)) + 
  egg::theme_article() + theme(legend.position = "none") + 
  xlab("Contribution to \u0394 wild bee abundance (landscape)")
svb.pol.margin <- sv_importance(ex.polmarg.glb, kind = "bee", max_display = 30,
                              viridis_args = list(option = "viridis"), size = 0.5,
                              show_numbers = TRUE, number_size = 2.7, format_fun = \(x) format(round(x,3),nsmall=3,)) +
  egg::theme_article() + theme(legend.position = "none") + 
  xlab("Contribution to \u0394 wild bee abundance (per margin)")

legend <- cowplot::get_legend(sv_importance(ex.polmarg.glb, kind = "bee", max_display = 5,
                              viridis_args = list(option = "viridis", name = "Metric value")) +
                                egg::theme_article() 
                              + theme(legend.box.margin = margin(0, 0, 0, 12)))

svb.pol.base.rename <- rename_fac_fun(svb.pol.base + theme(plot.margin = unit(c(1,0,0,0), "lines")))
svb.pol.delta.rename <- rename_fac_fun(svb.pol.delta + theme(plot.margin = unit(c(1,0,0,0), "lines")))
svb.pol.margin.rename <- rename_fac_fun(svb.pol.margin + theme(plot.margin = unit(c(1,0,0,0), "lines")))

imp.pol <- cowplot::plot_grid(
  svb.pol.base.rename, svb.pol.delta.rename, svb.pol.margin.rename, legend,
  nrow = 1, rel_widths = c(1,1,1,0.22),
  labels = c("(a)", "(b)", "(c)")
)

imp.pol

# ggsave("Figs/imp_pol.png", imp.pol, width = 12, height = 8, dpi = 300, bg = "white")
```

#### Nutrient export

```{r n_export_shaps, fig.width=12, fig.height=7}

svb.nexp.base <- sv_importance(ex.nexpbase.glb, kind = "bee", max_display = 30,
                              viridis_args = list(option = "viridis"), size = 0.5,
                              show_numbers = TRUE, number_size = 2.7, format_fun = \(x) format(round(x,3),nsmall=3,)) +
  egg::theme_article() + theme(legend.position = "none") + 
  xlab("Contribution to baseline nutrient runoff")
svb.nexp.delta <- sv_importance(ex.nexpdelta.glb, kind = "bee", max_display = 30,
                              viridis_args = list(option = "viridis"), size = 0.5,
                              show_numbers = TRUE, number_size = 2.7, format_fun = \(x) format(round(x,3),nsmall=3,)) + 
  egg::theme_article() + theme(legend.position = "none") + 
  xlab("Contribution to \u0394 nutrient runoff (landscape)")
svb.nexp.margin <- sv_importance(ex.nexpmarg.glb, kind = "bee", max_display = 30,
                              viridis_args = list(option = "viridis"), size = 0.5,
                              show_numbers = TRUE, number_size = 2.7, format_fun = \(x) format(round(x,3),nsmall=3,)) +
  egg::theme_article() + theme(legend.position = "none") + 
  xlab("Contribution to \u0394 nutrient runoff (per margin)")

svb.nexp.base.rename <- rename_fac_fun(svb.nexp.base + theme(plot.margin = unit(c(1,0,0,0), "lines")))
svb.nexp.delta.rename <- rename_fac_fun(svb.nexp.delta + theme(plot.margin = unit(c(1,0,0,0), "lines")))
svb.nexp.margin.rename <- rename_fac_fun(svb.nexp.margin + theme(plot.margin = unit(c(1,0,0,0), "lines")))

imp.nexp <- cowplot::plot_grid(
  svb.nexp.base.rename, svb.nexp.delta.rename, svb.nexp.margin.rename, legend,
  nrow = 1, rel_widths = c(1,1,1,0.22),
  labels = c("(a)", "(b)", "(c)")
)

imp.nexp
# ggsave("Figs/imp_nexp.png", imp.nexp, width = 12, height = 8, dpi = 300, bg = "white")
```

#### Sediment export

```{r sed_exp_shaps, fig.width=12, fig.height=7}

svb.sed.base <- sv_importance(ex.sedbase.glb, kind = "bee", max_display = 30,
                              viridis_args = list(option = "viridis"), size = 0.5,
                              show_numbers = TRUE, number_size = 2.7, format_fun = \(x) format(round(x,3),nsmall=3,)) +
  egg::theme_article() + theme(legend.position = "none") + 
  xlab("Contribution to baseline sediment runoff")
svb.sed.delta <- sv_importance(ex.seddelta.glb, kind = "bee", max_display = 30,
                              viridis_args = list(option = "viridis"), size = 0.5,
                              show_numbers = TRUE, number_size = 2.7, format_fun = \(x) format(round(x,3),nsmall=3,)) + 
  egg::theme_article() + theme(legend.position = "none") + 
  xlab("Contribution to \u0394 sediment runoff (landscape)")
svb.sed.margin <- sv_importance(ex.sedmarg.glb, kind = "bee", max_display = 30,
                              viridis_args = list(option = "viridis"), size = 0.5,
                              show_numbers = TRUE, number_size = 2.7, format_fun = \(x) format(round(x,3),nsmall=3,)) +
  egg::theme_article() + theme(legend.position = "none") + 
  xlab("Contribution to \u0394 sediment runoff (per margin)")

svb.sed.base.rename <- rename_fac_fun(svb.sed.base + theme(plot.margin = unit(c(1,0,0,0), "lines")))
svb.sed.delta.rename <- rename_fac_fun(svb.sed.delta + theme(plot.margin = unit(c(1,0,0,0), "lines")))
svb.sed.margin.rename <- rename_fac_fun(svb.sed.margin + theme(plot.margin = unit(c(1,0,0,0), "lines")))

imp.sed <- cowplot::plot_grid(
  svb.sed.base.rename, svb.sed.delta.rename, svb.sed.margin.rename, legend,
  nrow = 1, rel_widths = c(1,1,1,0.22),
  labels = c("(a)", "(b)", "(c)")
)

imp.sed
# ggsave("Figs/imp_sed.png", imp.sed, width = 12, height = 8, dpi = 300, bg = "white")

```

#### Baseflow

```{r qb_shaps, fig.width=12, fig.height=7}

svb.qb.base <- sv_importance(ex.qbbase.glb, kind = "bee", max_display = 30,
                              viridis_args = list(option = "viridis"), size = 0.5,
                              show_numbers = TRUE, number_size = 2.7, format_fun = \(x) format(round(x,3),nsmall=3,)) +
  egg::theme_article() + theme(legend.position = "none") + 
  xlab("Contribution to baseline groundwater recharge")
svb.qb.delta <- sv_importance(ex.qbdelta.glb, kind = "bee", max_display = 30,
                              viridis_args = list(option = "viridis"), size = 0.5,
                              show_numbers = TRUE, number_size = 2.7, format_fun = \(x) format(round(x,3),nsmall=3,)) + 
  egg::theme_article() + theme(legend.position = "none") + 
  xlab("Contribution to \u0394 groundwater recharge (landscape)")
svb.qb.margin <- sv_importance(ex.qbmarg.glb, kind = "bee", max_display = 30,
                              viridis_args = list(option = "viridis"), size = 0.5,
                              show_numbers = TRUE, number_size = 2.7, format_fun = \(x) format(round(x,3),nsmall=3,)) +
  egg::theme_article() + theme(legend.position = "none") + 
  xlab("Contribution to \u0394 groundwater recharge (per margin)")

svb.qb.base.rename <- rename_fac_fun(svb.qb.base + theme(plot.margin = unit(c(1,0,0,0), "lines")))
svb.qb.delta.rename <- rename_fac_fun(svb.qb.delta + theme(plot.margin = unit(c(1,0,0,0), "lines")))
svb.qb.margin.rename <- rename_fac_fun(svb.qb.margin + theme(plot.margin = unit(c(1,0,0,0), "lines")))

imp.qb <- cowplot::plot_grid(
  svb.qb.base.rename, svb.qb.delta.rename,svb.qb.margin.rename, legend,
  nrow = 1, rel_widths = c(1,1,1,0.22),
  labels = c("(a)", "(b)", "(c)")
)

imp.qb
# ggsave("Figs/imp_qb.png", imp.qb, width = 12, height = 8, dpi = 300, bg = "white")

```

### Across ecosystem services

```{r shap_fun}

# function to process the shap data

sv_fun <- function(shapviz_list, metric, scale = TRUE){
  svd.df <- lapply(shapviz_list, \(x) shapviz::sv_dependence(x, v=metric)$data) %>%
    setNames(names(shapviz_list)) %>% bind_rows(.id = "ES") %>%
    select(c("ES","shap",metric))
  
  if(scale){
    pred.sd <- lapply(shapviz_list, \(x) sd(apply(x$S, 1, sum))) %>%
    setNames(names(shapviz_list)) %>% bind_rows() %>%
      pivot_longer(cols = pollinators:n_export, names_to = "ES", values_to = "pred.sd")
    
    svd.df <- svd.df %>% left_join(pred.sd, by = "ES") %>%
      mutate(shap.scale = shap/pred.sd)
  }
  svd.df
}

# ES labeler 2
ES_labels2 <- c(
  pollinators = "Wild bee abundance",
  qb = "Groundwater recharge",
  sed_export = "Sediment runoff",
  n_export = "Nutrient runoff"
)

# base data
base.glb <- list(pollinators = ex.polbase.glb, qb = ex.qbbase.glb, 
                 sed_export = ex.sedbase.glb, n_export = ex.nexpbase.glb)

# delta data
delta.glb <- list(pollinators = ex.poldelta.glb, qb = ex.qbdelta.glb, 
                 sed_export = ex.seddelta.glb, n_export = ex.nexpdelta.glb)

# delta data
margin.glb <- list(pollinators = ex.polmarg.glb, qb = ex.qbmarg.glb, 
                 sed_export = ex.sedmarg.glb, n_export = ex.nexpmarg.glb)

```

#### Across percent natural landscape

```{r across_ntr, fig.height=3, fig.width=8}

shapval_ntr.base <- sv_fun(base.glb, "ntr_pland")
shapval_ntr.delta <- sv_fun(delta.glb, "ntr_pland")
shapval_ntr.marg <- sv_fun(margin.glb, "ntr_pland")

shapval_ntr.all <- list(base = shapval_ntr.base, delta = shapval_ntr.delta, delta.margin = shapval_ntr.marg) %>%
  bind_rows(.id = "response") %>%
  mutate(ES = factor(ES, levels = c("pollinators","qb", "sed_export", "n_export")))

shapd.ntrp <- ggplot(shapval_ntr.all, aes(ntr_pland, shap.scale, color = ES)) + 
  geom_hline(yintercept = 0, color = "gray30") +
  geom_point(size=0.5) + 
  geom_smooth(method = "gam", se = FALSE) +
  scale_color_viridis_d(option = "turbo", labels = ES_labels2, alpha = 0.5) + egg::theme_article() +
  facet_wrap(~response,# scales = "free", 
             labeller = as_labeller(response_type_labels)) +
  xlab("Natural habitat % landscape") + ylab("Scaled contribution\nto ES prediction") +
  theme(legend.position = "bottom", legend.title = element_blank())

shapd.ntrp

# ggsave("Figs/shapd_ntr_pland.png", shapd.ntrp, width = 8, height = 3, dpi = 300, bg = "white")


```

#### Across annual agriculture percent land cover

Important for nutrient export

```{r across_ann, fig.height=3, fig.width=8}

shapval_ann.base <- sv_fun(base.glb, "ann_agr_pland")
shapval_ann.delta <- sv_fun(delta.glb, "ann_agr_pland")
shapval_ann.marg <- sv_fun(margin.glb, "ann_agr_pland")

shapval_ann.all <- list(base = shapval_ann.base, 
                          delta = shapval_ann.delta, 
                          delta.margin = shapval_ann.marg) %>%
  bind_rows(.id = "response") %>%
  mutate(ES = factor(ES, levels = c("pollinators","qb", "sed_export", "n_export")))

annagr.importantvar <- shapval_ann.all %>%
  dplyr::filter(
    response == "delta" & ES == "n_export"
  )

annagr.notimportantvar <- shapval_ann.all %>%
  dplyr::filter(
    !(response == "delta" & ES == "n_export")
  )

annagr_legend <- ggplot(shapval_ann.all, 
                        aes(ann_agr_pland, shap.scale, color = ES)) + 
  geom_point(size=0.5) + 
  stat_smooth(geom = "line", method = "gam", se = FALSE) + 
  scale_color_viridis_d(option = "turbo", labels = ES_labels2, alpha = 0.5) + 
  egg::theme_article() +
  facet_wrap(~response, # scales = "free", 
             labeller = as_labeller(response_type_labels)) +
  theme(legend.position = "bottom", legend.title = element_blank()) +
  xlab("Annual agriculture % landscape") + 
  ylab("Scaled contribution\nto ES prediction") +
  guides(colour = guide_legend(label.vjust = 2))

annagr_legend <- annagr_legend %>% cowplot::get_legend()

shapd.annp <- ggplot(annagr.notimportantvar, aes(ann_agr_pland, shap.scale, color = ES)) + 
  geom_hline(yintercept = 0, color = "gray30") +
  geom_point(size=0.5, alpha = 0.1) + 
  stat_smooth(geom = "line", method = "gam", se = FALSE, alpha = 0.2) + 
  geom_point(data = annagr.importantvar, size=0.5, alpha = 0.6) +
  stat_smooth(data = annagr.importantvar, geom = "line", method = "gam", se = FALSE, alpha = 0.6) +
  scale_color_viridis_d(option = "turbo", labels = ES_labels2, alpha = 0.5) + 
  egg::theme_article() +
  facet_wrap(~response, # scales = "free", 
             labeller = as_labeller(response_type_labels)) +
  theme(legend.position = "none") +
  xlab("Annual agriculture % landscape") + 
  ylab("Scaled contribution\nto ES prediction")


shapd.annp.tag <- shapd.annp %>% egg::tag_facet() + 
  theme(strip.text = element_text(size = 8, margin = margin(r=1,b=6,l=6)))

shapd.annp.tag.legend <- cowplot::plot_grid(
  shapd.annp.tag, annagr_legend, ncol = 1, nrow = 2,
  rel_heights = c(10,1)
)
shapd.annp.tag.legend

# ggsave("Figs/SupFig_S6annagrpland.png", shapd.annp.tag.legend, width = 8, height = 3, dpi = 300, bg = "white")

```

#### Across annual agriculture edge density

Important for sediment export (and pollinator delta)

```{r across_annagr_ed, fig.height=3, fig.width=8}

shapval_anned.base <- sv_fun(base.glb, "ann_agr_ed", scale = TRUE)
shapval_anned.delta <- sv_fun(delta.glb, "ann_agr_ed", scale = TRUE)
shapval_anned.marg <- sv_fun(margin.glb, "ann_agr_ed", scale = TRUE)

shapval_anned.all <- list(base = shapval_anned.base, 
                          delta = shapval_anned.delta, 
                          delta.margin = shapval_anned.marg) %>%
  bind_rows(.id = "response") %>%
  mutate(ES = factor(ES, levels = c("pollinators","qb", "sed_export", "n_export")))

shapd.anned <- ggplot(shapval_anned.all, aes(ann_agr_ed, shap.scale, color = ES)) +
  geom_hline(yintercept = 0, color = "gray30") +
  geom_point(size=0.5) + 
  geom_smooth(method = "gam", se = FALSE) + 
  scale_color_viridis_d(option = "turbo", labels = ES_labels2, alpha = 0.5) + 
  egg::theme_article() +
  facet_wrap(~response, #scales = "free",
             labeller = as_labeller(response_type_labels)) + 
  theme(legend.position = "bottom", legend.title = element_blank()) + 
  xlab("Annual agriculture edge density") + ylab("Scaled contribution\nto ES prediction")
  

shapd.anned

# ggsave("Figs/shapd_ann_agr_ed.png", shapd.anned, width = 8, height = 3, dpi = 300, bg = "white")

```

#### Across annual agriculture cohesion

Important for baseflow

```{r across_annagr_cohesion, fig.height=3, fig.width=8}

shapval_anncoh.base <- sv_fun(base.glb, "ann_agr_cohesion")
shapval_anncoh.delta <- sv_fun(delta.glb, "ann_agr_cohesion")
shapval_anncoh.marg <- sv_fun(margin.glb, "ann_agr_cohesion")

shapval_anncoh.all <- list(base = shapval_anncoh.base, 
                          delta = shapval_anncoh.delta, 
                          delta.margin = shapval_anncoh.marg) %>%
  bind_rows(.id = "response") %>%
  mutate(ES = factor(ES, levels = c("pollinators","qb", "sed_export", "n_export")))

shapd.anncoh <- ggplot(shapval_anncoh.all, aes(ann_agr_cohesion, shap.scale, color = ES)) +
  geom_hline(yintercept = 0, color = "gray30") +
  geom_point(size=0.5) + 
  geom_smooth(method = "gam", se = FALSE) + 
  scale_color_viridis_d(option = "turbo", labels = ES_labels2, alpha = 0.5) + 
  egg::theme_article() +
  facet_wrap(~response, #scales = "free", 
             labeller = as_labeller(response_type_labels)) + 
  theme(legend.position = "bottom", legend.title = element_blank()) +
  xlab("Annual agriculture cohesion") + ylab("Scaled contribution\nto ES prediction")

shapd.anncoh

# ggsave("Figs/shapd_ann_agr_cohesion.png", shapd.anncoh, width = 8, height = 3, dpi = 300, bg = "white")

```

#### Across natural land cover edge density

Important for nutrient export

```{r across_ntr_ed, fig.height=3, fig.width=8}

shapval_ntred.base <- sv_fun(base.glb, "ntr_ed")
shapval_ntred.delta <- sv_fun(delta.glb, "ntr_ed")
shapval_ntred.marg <- sv_fun(margin.glb, "ntr_ed")

shapval_ntred.all <- list(base = shapval_ntred.base, 
                          delta = shapval_ntred.delta, 
                          delta.margin = shapval_ntred.marg) %>%
  bind_rows(.id = "response") %>%
  mutate(ES = factor(ES, levels = c("pollinators","qb", "sed_export", "n_export")))

shapd.ntred <- ggplot(shapval_ntred.all, aes(ntr_ed, shap.scale, color = ES)) +
  geom_hline(yintercept = 0, color = "gray30") +
  geom_point(size=0.5) + 
  geom_smooth(method = "gam", se = FALSE) + 
  scale_color_viridis_d(option = "turbo", labels = ES_labels2, alpha = 0.5) + 
  egg::theme_article() +
  facet_wrap(~response, # scales = "free",
             labeller = as_labeller(response_type_labels)) +
  theme(legend.position = "bottom", legend.title = element_blank()) +
  xlab("Natural habitat edge density") + ylab("Scaled contribution\nto ES prediction")

shapd.ntred

# ggsave("Figs/shapd_ntr_ed.png", shapd.ntred, width = 8, height = 3, dpi = 300, bg = "white")

```

### Combined plot

```{r combine_shap, fig.height=10, fig.width=6.5}

shared.legend <- cowplot::get_plot_component(shapd.ntred, 'guide-box-bottom', return_all = TRUE)

combine_shap <- cowplot::plot_grid(
  shapd.ntrp + theme(legend.position = "none"),
  shapd.anned + theme(legend.position = "none"),
  shapd.anncoh + theme(legend.position = "none"),
  shapd.ntred + theme(legend.position = "none"),
  shared.legend,
  ncol = 1, labels = c("(a)", "(b)", "(c)", "(d)"),
  rel_heights = c(1,1,1,1,0.1)
)

combine_shap

ggsave("Figs/shapd_combine.png", combine_shap, bg="white",height=10,width=6.5,units="in",dpi=300)


```

```{r facet_shap}

feature_labels <- c(
    ntr_pland = "% Natural land cover",
    ann_agr_ed = "Annual agriculture edge density",
    ann_agr_cohesion = "Annual agriculture cohesion",
    ntr_ed = "Natural land cover edge density"
)

shapd.combine <- list(
  ntr_pland = shapval_ntr.all %>% rename(feature.val = ntr_pland),
  ann_agr_ed = shapval_anned.all %>% rename(feature.val = ann_agr_ed),
  ann_agr_cohesion = shapval_anncoh.all %>% rename(feature.val = ann_agr_cohesion),
  ntr_ed = shapval_ntred.all %>% rename(feature.val = ntr_ed)
) %>% bind_rows(.id = "feature") %>%
  mutate(feature = factor(feature, levels = c("ntr_pland","ann_agr_ed","ann_agr_cohesion","ntr_ed")))

importantvar <- shapd.combine %>%
  dplyr::filter(
    (ES == "pollinators" & response == "base" & feature == "ntr_pland") |
    (ES == "pollinators" & response == "delta" & feature == "ann_agr_ed") |
    (ES == "pollinators" & response == "delta.margin" & feature == "ntr_pland") |
    (ES == "qb" & feature == "ann_agr_cohesion") |
    (ES == "sed_export" & feature == "ann_agr_ed") |
    (ES == "n_export" & response == "base" & feature == "ntr_pland") |
    (ES == "n_export" & response == "delta.margin" & feature == "ntr_ed")
  )

notimportantvar <- shapd.combine %>%
  dplyr::filter(
    !((ES == "pollinators" & response == "base" & feature == "ntr_pland") |
       (ES == "pollinators" & response == "delta" & feature == "ann_agr_ed") |
       (ES == "pollinators" & response == "delta.margin" & feature == "ntr_pland") |
       (ES == "qb" & feature == "ann_agr_cohesion") |
       (ES == "sed_export" & feature == "ann_agr_ed") |
       (ES == "n_export" & response == "base" & feature == "ntr_pland") |
       (ES == "n_export" & response == "delta.margin" & feature == "ntr_ed"))
  )


shapd.facet <- ggplot(notimportantvar, aes(feature.val, shap.scale, color = ES)) +
  geom_hline(yintercept = 0, color = "gray30") +
  geom_point(size=0.5, alpha = 0.1) + 
  stat_smooth(geom = "line", method = "gam", se = FALSE, alpha = 0.2) + 
  geom_point(data = importantvar, size=0.5, alpha = 0.6) +
  stat_smooth(data = importantvar, geom = "line", method = "gam", se = FALSE, alpha = 0.6) +
  scale_color_viridis_d(option = "turbo", labels = ES_labels2) + 
  egg::theme_article() +
  ggh4x::facet_grid2(response~feature, scales = "free", #independent = "y",
                     switch = "both",
             labeller = labeller(response=response_type_labels, feature = feature_labels)) +
  # facet_grid(response~feature, scales = "free_x",
  #            labeller = as_labeller(response_type_labels)) +
  theme(legend.position = "bottom", legend.title = element_blank()) +
  xlab("Predictor value") + ylab("Scaled contribution to ES prediction") +
  theme(strip.placement = "outside") +
  guides(colour = guide_legend(label.vjust = 2))

shapd.facet.tag <- shapd.facet %>% egg::tag_facet() + 
  theme(strip.text = element_text(size = 8, margin = margin(r=1,b=6,l=6)))
shapd.facet.tag

ggsave("Figs/shapd_combine_facets.png", shapd.facet.tag, bg="white",height=5.5,width=8,units="in",dpi=300)

```


```{r extra}

# fit some gams
dat.rf.gam <- shapd.combine %>% 
  group_by(response, feature, ES) %>%
  nest() %>%
  # Fit GAM
  mutate(Mod = map(data, ~mgcv::gam(shap.scale ~ s(feature.val), data = .x))) %>%
  # Get R2
  mutate(R2 = map_dbl(Mod, ~round(summary(.x)$r.sq, 2)),
         mae = map_dbl(Mod, ~round(mae.fun(.x), 3))) %>%
  mutate(
    edf =  map_dbl(Mod, ~round(k.check.fun(.x)[2], 3)),
    kindex =  map_dbl(Mod, ~round(k.check.fun(.x)[3], 3)),
    pvalue =  map_dbl(Mod, ~round(k.check.fun(.x)[4], 3))
  )

test <- ggeffects::ggpredict(dat.rf.gam$Mod[[48]])

test2 <- ggeffects::ggpredict(dat.gam$Mod[[5]])

ggplot(shapd.combine %>% filter(feature == "ann_agr_ed" & response == "delta" & ES == "pollinators"),
       aes(feature.val, shap.scale)) + geom_point() + geom_smooth(method = "gam") + 
  geom_line(data = test[[1]], mapping = aes(x, predicted), inherit.aes = FALSE)

corplot_ntrpland_annagred <- ggplot(dat, aes(ntr_pland,ann_agr_ed)) + 
  geom_point(size = 0.5) + 
  xlab("% Natural land cover") + ylab("Annual agriculture edge density") +
  geom_label(x = Inf, y = Inf,
            label = paste("r = ", round(cor(dat$ntr_pland, dat$ann_agr_ed),2), sep = " "),
            hjust = 1, vjust = 1) +
  egg::theme_article()

ggplot(dat, aes(ann_agr_ed, delta_pollinators)) + geom_point()
ggplot(dat, aes(ntr_pland,delta_pollinators)) + geom_point()

# ggsave("Figs/SupFig_ntrpland_annagred.png", corplot_ntrpland_annagred, height = 3, width = 4, units = "in", bg = "white", dpi = 300)




```